Strong lensing selection effects

### 2022-07-24 paper/seleff.tex
constraining the scatter parameters

We drew stellar masses from the stellar mass function of quiescent galaxies measured by \citet{Muz++13}.
In particular, we used the following comoving number density distribution
\begin{equation}
\Phi(\mobs) = (\ln{10})\Phi^*\left[10^{(\mobs - \mstar^*)(1+\alpha)}\right]\times \exp{\left[-10^{(\mobs -\mstar^*)}\right]}.
\end{equation}
We set $\Phi^*=1.009\times10^{-3}\,{\rm Mpc}^{-3}$, $\alpha=-0.92$ and $\log{\mstar^*}=11.21$. These are the best-fit values measured by \citet{Muz++13}.

@@

We then proceeded to fit for the axis ratio distribution of the same sample of galaxies.
\Fref{fig:qhist} shows a histogram of the observed distribution.
We fitted this with a beta disribution:
\begin{equation}\label{eq:betaappendix}
\pr(q) \propto q^{\alpha-1}(1 - q)^{\beta-1}.
\end{equation}
We obtained $\alpha=6.28$ and $\beta=2.05$.
\begin{figure}
\includegraphics[width=\columnwidth]{q_dist.eps}
\caption{
Distribution in axis ratio of a sample of early-type galaxies.
The sample consists of $8078$ galaxies from the SDSS, selected by means of cuts in redshift, colour and S\'{e}rsic index as explained in the text.
Measurements of the axis ratio are taken from the de Vaucouleurs model fits of \citet{Mee++15}.
The model curve is a beta distribution, \Eref{eq:betaappendix}, with $\alpha=6.28$ and $\beta=2.05$.
}
\end{figure}

@@

We used the fundamental plane of early-type galaxies to set an upper limit on the intrinsic scatter parameters of the simulation: $\sigma_{\mathrm{sps}}$, $\sigma_h$ and $\sigma_\gamma$.
First, we measured the fundamental plane of the sample of SDSS early-type galaxies introduced in Appendix~\ref{sect:appendixa}.
We took measurements of the line-of-sight stellar velocity dispersion within the SDSS spectroscopic aperture, $\sigma_{\mathrm{ap}}$, from the SDSS data release 16 catalogue \citep{Ahu++20}.
Then, we fitted the following model to the distribution of $\sigma_{\mathrm{ap}}$ of the sample:
\begin{equation}
\pr(\sigma_{\mathrm{ap}}) \sim \mathcal{N}(\mu_{\sigma} + \beta(\log{\mobs} - 11.4) + \xi(\log{\reff} - 1.2), \sigma_{\sigma}).
\end{equation}
We accounted for observational uncertainties on $\sigma_{\mathrm{ap}}$ when doing the fit, therefore the parameter $\sigma_{\sigma}$ describes the intrinsic scatter in the logarithm of the velocity dispersion, deconvolved from the observational scatter.
In principle we should also account for observational uncertainties on the stellar mass measurement, as they too contribute to the inferred scatter in $\sigma_{\mathrm{ap}}$. In practice, however, it is difficult to estimate observational uncertainties on stellar population synthesis measurements.
For this reason we chose not to propagate uncertainties on $\mobs$. As a result, the inferred scatter parameter $\sigma_{\sigma}$ is slightly overestimated.
We obtained $\mu_{\sigma}=2.36$, $\beta_{\sigma}=0.33$, $\xi_\sigma=-0.17$, and $\sigma_{\sigma}=0.035$.

We then generated samples of galaxies from the model of section \ref{ssec:lenses}.

We did not attempt to match the other parameters of the fundamental plane (i.e. the constant of proportionality of \Eref{eq:fpcompact} and the power-law indices $\beta_{\sigma}$ and $\xi_{\sigma}$), because these are sensitive to the orbital anisotropy of the galaxies, which we are asserting to be zero.

%First, we took the sample of SDSS early-type galaxies introduced in Appendix~\ref{sect:appendixa}

>>>>

We drew stellar masses from the stellar mass function of quiescent galaxies measured by \citet{Muz++13}.
In particular, we used the following comoving number density distribution
\begin{equation}
\Phi(\mobs) = \Phi^*\left(\frac{\mobs}{\mstar^*}\right)^{\alpha} \exp{\left\{-\frac{\mobs}{\mstar^*}\right\}}.
\end{equation}
We set $\Phi^*=1.009\times10^{-3}\,{\rm Mpc}^{-3}$, $\alpha=-0.92$ and $\log{\mstar^*}=11.21$: these are the best-fit values measured by \citet{Muz++13}.

@@

We then proceeded to fit for the axis ratio distribution of the same sample of galaxies.
\Fref{fig:qhist} shows a histogram of the observed distribution.
We fitted this with a beta disribution:
\begin{equation}\label{eq:betaappendix}
\pr(q) \propto q^{\alpha-1}(1 - q)^{\beta-1}.
\end{equation}
We obtained $\alpha=6.28$ and $\beta=2.05$.
\begin{figure}
\includegraphics[width=\columnwidth]{q_dist.eps}
\caption{
Distribution in axis ratio of a sample of early-type galaxies.
The sample consists of $8078$ galaxies from the SDSS, selected by means of cuts in redshift, colour and S\'{e}rsic index as explained in the text.
Measurements of the axis ratio are taken from the de Vaucouleurs model fits of \citet{Mee++15}.
The model curve is a beta distribution, \Eref{eq:betaappendix}, with $\alpha=6.28$ and $\beta=2.05$.
\label{fig:qhist}
}
\end{figure}

@@

We used the fundamental plane of early-type galaxies to set an upper limit on the intrinsic scatter parameters of the simulation: $\sigma_{\mathrm{sps}}$, $\sigma_h$ and $\sigma_\gamma$.
First, we measured the fundamental plane of the sample of SDSS early-type galaxies introduced in Appendix~\ref{sect:appendixa}.
We took measurements of the line-of-sight stellar velocity dispersion within the SDSS spectroscopic aperture, $\sigma_{\mathrm{ap}}$, from the SDSS data release 16 catalogue \citep{Ahu++20}.
Then, we fitted the following model to the distribution of $\sigma_{\mathrm{ap}}$ of the sample:
\begin{equation}\label{eq:fpfull}
\pr(\sigma_{\mathrm{ap}}) \sim \mathcal{N}(\mu_{\sigma} + \beta(\log{\mobs} - 11.4) + \xi(\log{\reff} - 1.2), \sigma_{\sigma}).
\end{equation}
We accounted for observational uncertainties on $\sigma_{\mathrm{ap}}$ when doing the fit, therefore the parameter $\sigma_{\sigma}$ describes the intrinsic scatter in the logarithm of the velocity dispersion, deconvolved from the observational scatter.
In principle we should also account for observational uncertainties on the stellar mass measurement, as they too contribute to the inferred scatter in $\sigma_{\mathrm{ap}}$. In practice, however, it is difficult to estimate observational uncertainties on stellar population synthesis measurements.
For this reason we chose not to propagate uncertainties on $\mobs$. As a result, the inferred scatter parameter $\sigma_{\sigma}$ is slightly overestimated.
We obtained $\mu_{\sigma}=2.36$, $\beta_{\sigma}=0.33$, $\xi_\sigma=-0.17$, and $\sigma_{\sigma}=0.035$.

We then generated samples of $z=0.2$ galaxies from the model of section \ref{ssec:lenses} and used the spherical Jeans equation to predict their central velocity dispersion.
The spherical Jeans equation is \citep{B+T87}
\begin{equation}\label{eq:jeans}
\frac{d(\rho_*\sigma_r^2)}{dr} + \frac{\beta(r)}{r}\rho_*\sigma_r^2 = -\rho_*(r)\frac{GM(r)}{r^2},
\end{equation}
where $\rho_*(r)$ is the 3-dimensional distribution of dynamical tracers (i.e. the stars), $\sigma_r$ the radial component of the velocity dispersion, $\beta(r)$ the orbital anisotropy parameter, and $M(r)$ the mass enclosed within a spherical shell of radius $r$. 
We assumed isotropic orbits ($\beta=0$), then integrated \Eref{eq:jeans} to obtain the seeing-convolved, surface brightness-weighted line-of-sight velocity dispersion within the SDSS spectroscopic aperture\footnote{We used Python code available at \url{https://github.com/astrosonnen/spherical_jeans}, for this purpose.}.
Finally, we fitted the fundamental plane relation of \Eref{eq:fpfull} to the observed stellar mass, half-light radius and central velocity dispersion of the mock sample.
We repeated this procedure for different values of the intrinsic scatter parameters $\sigma_{\mathrm{sps}}$, $\sigma_\mathrm{h}$ and $\sigma_\gamma$.

%We did not attempt to match the other parameters of the fundamental plane (i.e. the constant of proportionality of \Eref{eq:fpcompact} and the power-law indices $\beta_{\sigma}$ and $\xi_{\sigma}$), because these are sensitive to the orbital anisotropy of the galaxies, which we are asserting to be zero.
